# Manual chill analysis

The data that is used in this lesson is from the inbuilt `Winters_hours_gaps` dataset. It was recorded 2008 in a walnut orchard in Winters, California. The dataset was modified by excluding a not used `Temp_gaps` column.

Table for the modified `Winters_hours_gaps` dataset

```{r, echo=FALSE}
require(chillR)
hourtemps <- subset(Winters_hours_gaps, select = -c(Temp_gaps))

row.names(hourtemps) <- NULL
require(kableExtra)
require(dplyr)
kbl(hourtemps[1:5,], escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")
```

**1.Write a basic function that calculates warm hours (\>25°C)**

```{r}
WH <- function(hourtemps)
  {
  hourtemps[,"Warming_hours"] <- hourtemps$Temp>25
  return(hourtemps)
}
hourtemps_plus_WH <- WH(hourtemps)
```

\

**2.Apply this function to the Winters_hours_gaps dataset**

Table for `hourtemps_plus_WH`

```{r, echo=FALSE}
hourtemps_plus_WH <- hourtemps_plus_WH[4008:4018, 1:5]
row.names(hourtemps_plus_WH) <- NULL
hourtemps_plus_WH$Temp <- ifelse(
  hourtemps_plus_WH$Temp > 25,
  cell_spec(hourtemps_plus_WH$Temp, color = "red", bold = T),
  cell_spec(hourtemps_plus_WH$Temp, color = "green")
)

kbl(hourtemps_plus_WH, escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")
```

**3.Extend this function, so that it can take start and end dates as inputs and sums up warm hours between these dates**

```{r}
WH_YEARMODA <- function(hourtemps, start_dateYEARMODA, end_dateYEARMODA, start_hour=12, end_hour=12)
{

  start_year <- as.numeric(substr(start_dateYEARMODA, 1, 4))
  start_month <- as.numeric(substr(start_dateYEARMODA, 5, 6))
  start_day <- as.numeric(substr(start_dateYEARMODA, 7, 8))

  end_year <- as.numeric(substr(end_dateYEARMODA, 1, 4))
  end_month <- as.numeric(substr(end_dateYEARMODA, 5, 6))
  end_day <- as.numeric(substr(end_dateYEARMODA, 7, 8))
  

  start_date <- which(hourtemps$Year==start_year & hourtemps$Month==start_month
                      & hourtemps$Day==start_day & hourtemps$Hour==start_hour)
  end_date <- which(hourtemps$Year==end_year & hourtemps$Month==end_month
                    & hourtemps$Day==end_day & hourtemps$Hour==end_hour)
  
  warm_hours <- WH(hourtemps)
  
  WHs <- sum(warm_hours[start_date:end_date, "Warming_hours"])
  return(WHs)
}

warming_hours_for_time_frame <- WH_YEARMODA(hourtemps, 20080501, 20080831)

```

Between the 1th of May 2008 and the 31th of August 2008, **`r warming_hours_for_time_frame`** hours over 25°C occured.

------------------------------------------------------------------------