# Chill Models

## Run the chilling() function on the Winters_hours_gap dataset

The chilling function calculates different chilling methods and GDH.

```{r 07 running chilling function, message=FALSE, warning=FALSE}
require(chillR)
chilling_output <- chilling(make_JDay(Winters_hours_gaps),Start_JDay = 80, End_JDay = 120)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
require(kableExtra)
require(dplyr)
title <- "Chilling output for a part of the `Winters_hours_gap`"
chilling_output %>% 
kbl(escape = F, format = "html", caption = title) %>%
  kable_paper("hover", full_width = T, position = "left")
```

## Create your own temperature-weighting chill model using the step_model() function.

```{r 07 creating own step_model() function}
own_utah_steps <- data.frame(
  lower=c(-291,0,1,2,3,4,5,6),
  upper=c(0,1,2,3,4,5,6,100),
  weight=c(0,0.5,1,2,3,2,1,0))
```

```{r, echo=FALSE}
title <- "Table for `own_utah_steps`"
own_utah_steps %>% 
kbl(escape = F, format = "html", caption = title) %>%
  kable_paper("hover", full_width = T, position = "left")
```

Testing the modified `own_utah_steps` alone.
```{r 07 applying own step_model() function}
own_utah <- function(x) step_model(x,own_utah_steps)
test <- own_utah(Winters_hours_gaps$Temp)

head(test)
```

## Run this model on the Winters_hours_gaps dataset using the tempResponse() function.

In the next code chunk, the `Modified_Utah_Model` is included in the list of the `tempResponse()` function.
```{r}
output<-tempResponse(make_JDay(Winters_hours_gaps),
                     Start_JDay = 80, End_JDay = 100, 
                     models=list(Modified_Utah_Model=own_utah,Chill_portions=Dynamic_Model),
                     whole_record = FALSE)
```

```{r, echo=FALSE}
title <- "TempResponse output for the chill of the `Modified_Utah_Model` and `Chill_portions`"
output %>% 
kbl(escape = F, format = "html", caption = title) %>%
  kable_paper("hover", full_width = T, position = "left")
```

\
<hr />