# Making hourly temperatures

**1. Choose a location of interest, find out its latitude and produce plots of daily sunrise, sunset and daylength**

The chosen location with the latitude of '32' can be seen below in the leaflet. It is a farm located in the south of Morocco.
```{r, echo=FALSE, message=FALSE}
require(leaflet) 
m <- leaflet() %>% setView(lng = -8, lat = 32, zoom = 16)
m %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addMiniMap(
    tiles = providers$Esri.WorldStreetMap,
    toggleDisplay = TRUE)
  
```
\
The `chillR` `daylength` function can be used to get the data for the sunsets and sunrises for a specified latitude. The `melt` command of the `reshape2` package is used to bring the dataset in the "long" format, that is more suitable for the use with `ggplot2`.
```{r, message=FALSE}
require(chillR)
require(reshape2)

day_infos<-daylength(latitude=32,JDay=1:365)

days_df<-data.frame(JDay=1:365,
                    Sunrise=day_infos$Sunrise,
                    Sunset=day_infos$Sunset,
                    Daylength=day_infos$Daylength)
days_df<-melt(days_df, id=c("JDay"))
```

The `days_df` data is plotted in 3 line graphs (`geom_line`) split by the factor `variable`.
```{r, message=FALSE}
require(ggplot2)
ggplot(days_df, aes(JDay, value)) +
  geom_line(lwd=1.5) + facet_grid(cols=vars(variable)) +
  ylab("Time of Day / Daylength (Hours)") +
  theme_bw(base_size = 20)
```
\

**2. Produce an hourly dataset, based on idealized daily curves, for the KA_weather dataset (included in chillR)**

The `stack_hourly_temps` function is able to simulate idealized hourly data with daily minimal temperatures, daily maximal temperatures and the latitude of the place of interest.

```{r}
hourly_temps_KA <- stack_hourly_temps(KA_weather, latitude=50.4)

#str(hourly_temps_KA) # to get the correct item from the list

hourly_temps_KA_df <- hourly_temps_KA$hourtemps


hourly_temps_KA_df[,"connected_timedate"] <- ISOdate(
                                  year = hourly_temps_KA_df$Year,
                                  month = hourly_temps_KA_df$Month,
                                  day = hourly_temps_KA_df$Day,
                                  hour = hourly_temps_KA_df$Hour)

ggplot(hourly_temps_KA_df[3500:3660,], aes(connected_timedate,Temp)) +
  geom_line(lwd=1.5) +
  labs(x="Date",
       y="Temperature in C°",
       title = "Hourly modelled temperature for Week 22 in 2002") +
  theme_bw(base_size = 14)
```

**3. Produce empirical temperature curve parameters for the Winters_hours_gaps dataset, and use them to predict hourly values from daily temperatures (this is very similar to the example above, but please make sure you understand what’s going on)**

The `Empirical_daily_temperature_curve` function is used to extract hourly coefficients for each month from the `Winters_hours_gaps`. The `make_all_day_table` function is used to calculate `Tmin` and `Tmax` values, that are used in combination with the coefficients in the `Empirical_hourly_temperatures` to calculate the hourly temperatures.

```{r Calculating the empirical model, warning=FALSE, message=FALSE}
coeffs<-Empirical_daily_temperature_curve(Winters_hours_gaps)
winters_daily<-make_all_day_table(Winters_hours_gaps, input_timestep="hour")
winters_empirical<-Empirical_hourly_temperatures(winters_daily,coeffs)
```


```{r, warning=FALSE, message=FALSE, echo=FALSE}
require(dplyr)
require(kableExtra)
title <- "Modified Table for empirical hourly temperatures"
select(winters_empirical[1:10,], DATE, Tmin, Tmax, JDay, Hour, Temp) %>% 
kbl(escape = F, format = "html", caption = title) %>%
  kable_paper("hover", full_width = T, position = "left")

```
Hourly data with the Empirical and ideal method are compared with the real measured temperature. The `melt` function is again used to bring the dataset into the right shape.
```{r}
require(reshape2)

winters_empirical<-winters_empirical[,c("Year","Month","Day","Hour","Temp")]
colnames(winters_empirical)[ncol(winters_empirical)]<-"Temp_empirical"

winters_ideal<-stack_hourly_temps(winters_daily, latitude=38.5)$hourtemps
winters_ideal<-winters_ideal[,c("Year","Month","Day","Hour","Temp")]
colnames(winters_ideal)[ncol(winters_ideal)]<-"Temp_ideal"

winters_temps<-merge(Winters_hours_gaps,winters_empirical,
                     by=c("Year","Month","Day","Hour"))
winters_temps<-merge(winters_temps,winters_ideal,
                     by=c("Year","Month","Day","Hour"))

winters_temps[,"DATE"]<-ISOdate(winters_temps$Year,
                                winters_temps$Month,
                                winters_temps$Day,
                                winters_temps$Hour)

winters_temps_to_plot<-winters_temps[,c("DATE",
                                        "Temp",
                                        "Temp_empirical",
                                        "Temp_ideal")]
winters_temps_to_plot<-winters_temps_to_plot[200:262,]
winters_temps_to_plot<-melt(winters_temps_to_plot, id=c("DATE")) 
colnames(winters_temps_to_plot)<-c("DATE","Method","Temperature")


ggplot(data=winters_temps_to_plot, aes(DATE,Temperature, colour=Method)) +
  geom_line(lwd=1.3) +
  ylab("Temperature (°C)") +
  xlab("Date") +
  ggtitle ("Different temperature methods for Winters dataset") +
  scale_color_manual(labels = c("Observed", "Ideal", "Empirical"),
                     name="Temperature",
                     values = c("slateblue", "peru", "mediumaquamarine")) +
  theme_bw(base_size = 14)
```

------------------------------------------------------------------------