# Making hourly temperatures

**1.Choose a location of interest, find out its latitude and produce plots of daily sunrise, sunset and daylength**

The chosen location with the latitude of '32' can be seen below in the leaflet.
```{r, echo=FALSE, message=FALSE}
require(leaflet) 
m <- leaflet() %>% setView(lng = -8, lat = 32, zoom = 16)
m %>% addProviderTiles(providers$Esri.WorldImagery) %>%
  addMiniMap(
    tiles = providers$Esri.WorldStreetMap,
    toggleDisplay = TRUE)
  
```

```{r, message=FALSE}
require(chillR)
require(ggplot2)
require(reshape2)
day_infos<-daylength(latitude=32,JDay=1:365)

days_df<-data.frame(JDay=1:365,Sunrise=day_infos$Sunrise,Sunset=day_infos$Sunset,Daylength=day_infos$Daylength)
days_df<-melt(days_df, id=c("JDay"))

ggplot(days_df, aes(JDay, value)) + geom_line(lwd=1.5) + facet_grid(cols=vars(variable)) +
  ylab("Time of Day / Daylength (Hours)") + theme_bw(base_size = 20)
```


**2.Produce an hourly dataset, based on idealized daily curves, for the KA_weather dataset (included in chillR)**
```{r}

hourly_temps_KA <- stack_hourly_temps(KA_weather, latitude=50.4)

#str(hourly_temps_KA) # to get the correct item from the list

hourly_temps_KA_df <- hourly_temps_KA$hourtemps


hourly_temps_KA_df[,"connected_timedate"] <- ISOdate(year = hourly_temps_KA_df$Year,
                                  month = hourly_temps_KA_df$Month,
                                  day = hourly_temps_KA_df$Day,
                                  hour = hourly_temps_KA_df$Hour)

ggplot(hourly_temps_KA_df[3500:3660,], aes(connected_timedate,Temp)) +
  geom_line(lwd=1.5) +
  labs(x="Date", y="Temperature in C°", title = "Hourly modelled temperature for Week 22 in 2002") +
  theme_bw(base_size = 14)
```

**3.Produce empirical temperature curve parameters for the Winters_hours_gaps dataset, and use them to predict hourly values from daily temperatures (this is very similar to the example above, but please make sure you understand what’s going on)**

```{r}
require(dplyr)

kbl(Winters_hours_gaps[1:10,], escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")

coeffs<-Empirical_daily_temperature_curve(Winters_hours_gaps)

kbl(coeffs[1:10,], escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")

winters_daily<-make_all_day_table(Winters_hours_gaps, input_timestep="hour")

kbl(winters_daily[1:10,], escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")

winters_empirical<-Empirical_hourly_temperatures(winters_daily,coeffs)

kbl(winters_empirical[1:10,], escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")

```

```{r}
require(reshape2)

winters_empirical<-winters_empirical[,c("Year","Month","Day","Hour","Temp")]
colnames(winters_empirical)[ncol(winters_empirical)]<-"Temp_empirical"

winters_ideal<-stack_hourly_temps(winters_daily, latitude=38.5)$hourtemps
winters_ideal<-winters_ideal[,c("Year","Month","Day","Hour","Temp")]
colnames(winters_ideal)[ncol(winters_ideal)]<-"Temp_ideal"

winters_temps<-merge(Winters_hours_gaps,winters_empirical, by=c("Year","Month","Day","Hour"))
winters_temps<-merge(winters_temps,winters_ideal, by=c("Year","Month","Day","Hour"))

winters_temps[,"DATE"]<-ISOdate(winters_temps$Year,winters_temps$Month, winters_temps$Day, winters_temps$Hour)

winters_temps_to_plot<-winters_temps[,c("DATE","Temp","Temp_empirical","Temp_ideal")]
winters_temps_to_plot<-winters_temps_to_plot[200:262,]
winters_temps_to_plot<-melt(winters_temps_to_plot, id=c("DATE")) 
colnames(winters_temps_to_plot)<-c("DATE","Method","Temperature")


ggplot(data=winters_temps_to_plot, aes(DATE,Temperature, colour=Method)) +
  geom_line(lwd=1.3) +
  ylab("Temperature (°C)") +
  xlab("Date") +
  ggtitle ("Different temperature methods for Winters dataset") +
  scale_color_manual(labels = c("Observed", "Ideal", "Empirical"), name="Temperature", values = c("slateblue", "peru", "mediumaquamarine")) +
  theme_bw(base_size = 14)


```

------------------------------------------------------------------------