# Filling gaps in temperature records

## **1. Use chillR functions to find out how many gaps you have in this dataset (even if you have none, please still follow all further steps)**

```{r Load data and create artificial gap, echo=FALSE, message=FALSE, warning=FALSE}
cleaned_weather <- read.csv("data/CapeTown_chillR_weather.csv")
CapeTown_weather_gap <- cleaned_weather[-c(11700:12249),]
coords <- c(18.8148, -33.9575)
```

```{r Test for gaps in data}
require(chillR)
fixed_weather <- fix_weather(CapeTown_weather_gap)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
require(dplyr)
require(kableExtra)
kbl(fixed_weather$QC, escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")
```

**2. Create a list of the 25 closest weather stations using the handle_gsod function**
I created a new `patching_station_list` for the time interval from 2004 to 2006
```{r Command for creating the weather stations list, eval=TRUE}
patching_station_list<-handle_gsod(action="list_stations",
                          location=c(coords[1], coords[2]),
                          time_interval=c(2008,2011),
                          stations_to_choose_from = 25)
```

```{r, eval=FALSE, echo=FALSE}
write.csv(patching_station_list,"data/patching_station_list.csv",row.names=FALSE)
```


```{r Loading the station list and creating a table, echo=FALSE}
patching_station_list <- read.csv("data/patching_station_list.csv")

kbl(patching_station_list[1:10,], escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")
str(patching_station_list$STATION.NAME[1])
```



**3. Identify suitable weather stations for patching gaps**\
The ``r patching_station_list$STATION.NAME[2]``' station on the second position or the ``r patching_station_list$STATION.NAME[4]`` station on position 4 could be suitable stations for patching gaps in our main data set. The ``r patching_station_list$STATION.NAME[2]`` station is characterized by a close distance to the location of interest and the ``r patching_station_list$STATION.NAME[4]`` as a station that is less affected by coastal climate than other closer stations. ``r patching_station_list$STATION.NAME[6]`` serves as a last backup.

**4. Download weather data for promising stations, convert them to chillR format and compile them in a list**

```{r Looping through the data for patching, eval=FALSE}
positions_in_station_list<-c(2,4,6)
patch_weather<-list()

for(i in 1:length(positions_in_station_list))
   {patch_weather[[i]]<-handle_gsod(handle_gsod(action="download_weather",
                                    location=patching_station_list$chillR_code[
                                      positions_in_station_list[i]],
                                    time_interval=c(1977,2020)))[[1]]$weather
   names(patch_weather)[i]<-patching_station_list$STATION.NAME[
     positions_in_station_list[i]]
   }
```

```{r Saving the command for eventual changes, echo=FALSE, eval=FALSE}
save_temperature_scenarios(patch_weather,"data/", "patch_weather")
```

```{r Loading the saved data, echo=FALSE}
patch_weather <- load_temperature_scenarios("data/", "patch_weather")
```

**5. Use the patch_daily_temperatures function to fill gaps**

```{r Patching the main datset with the newly downloaded data, eval=TRUE}
patched <- patch_daily_temperatures(weather = CapeTown_weather_gap,
                                  patch_weather = patch_weather)
#                                 ,max_mean_bias = 1,
#                                  max_stdev_bias = 2)
```

The tables show the progress made during patching
```{r Creating table for patching stats, echo=FALSE}
kable(patched$statistics[[1]],
      caption=paste("Patch statistics for",names(patched$statistics)[1])) %>%
    kable_styling("hover", full_width = F, position = "left", font_size = 10)

 kable(patched$statistics[[2]],
       caption=paste("Patch statistics for",names(patched$statistics)[2])) %>%
     kable_styling("hover", full_width = F, position = "left", font_size = 10)
 
kable(patched$statistics[[3]],
       caption=paste("Patch statistics for",names(patched$statistics)[2])) %>%
     kable_styling("hover", full_width = F, position = "left", font_size = 10)
```

Stdev_bias too high!

**6. Investigate the results - have all gaps been filled?**

```{r Patching last gaps with fix_weather}
patched_fixed_weather <- fix_weather(patched)

final_patched_fixed_weather <- fix_weather(patched_fixed_weather)

```

```{r Creating tables for QC, echo=FALSE}
kable(patched_fixed_weather$QC,
      caption="Quality check after the patching procedure") %>%
  kable_styling("hover", position = "left", font_size = 10)

kable(summary(final_patched_fixed_weather$QC),
      caption="Quality check after fixing left over gaps") %>%
  kable_styling("hover", position = "left", font_size = 10)

```

**7. If necessary, repeat until you have a dataset you can work with in further analyses**
The last single day gaps were fixed with the `fixed_weather` command. For further analyses, i will use the original dataset without the artificial gap. Gaps of single days are interpolated with `fixed_weather`.

```{r Loading original data and fixing single missing days, eval=FALSE}
cleaned_weather <- read.csv("data/CapeTown_chillR_weather.csv")
```

```{r Data saving, echo=FALSE, eval=FALSE}
write.csv(fixed_weather$weather, "data/CapeTown_weather.csv", row.names=FALSE)
```