# Generating temperature scenarios

## For the location you chose for your earlier analyses, use chillR’s weather generator to produce 100 years of synthetic temperature data.

```{r 11 Loading and adjusting needed data, echo=FALSE, message=FALSE, warning=FALSE}
require(chillR)
require(ggplot2)
coords <- c(18.8148, -33.9575)
weather_CapeTown <- read.csv("data/CapeTown_weather.csv")
```

We are using the `temperature_generation` function of the `chillR` package to generate 100 years of temperature that is calibrated on our input data `weather_CapeTown`.

```{r 11 Using the temperature_generation to create weather scenarios, eval=FALSE, message=FALSE}
generated_temp<-temperature_generation(weather_CapeTown,
                             years=c(1990,2005),
                             sim_years = c(2001,2100))
```

```{r 11 Saving the generated weather, eval=FALSE, echo=FALSE}
write.csv(generated_temp, "data/generated_temp.csv", row.names=FALSE)
```

```{r 11 Loading data from weather generation, echo=FALSE, message=FALSE}
generated_temp <- read.csv("data/generated_temp.csv")
```

In the following chunk of code, we transformed and combined our datasets into one dataset that has the most suitable shape for the use with the `ggplot` function.

```{r 11 Transforming data from weather generation, echo=TRUE, message=FALSE}
temperatures<-cbind(weather_CapeTown[
  which(weather_CapeTown$Year %in% 1990:2005),], Data_source="observed")

temperatures<-rbind(temperatures[,c("Year","Month","Day","Tmin","Tmax","Data_source")],
                    cbind(generated_temp[,c("Year","Month","Day","Tmin","Tmax")],
                          Data_source="simulated"))

temperatures[,"Date"]<-as.Date(ISOdate(2000,
                                       temperatures$Month,
                                       temperatures$Day))
```

We plot the two different `Data_source`s in two plots for a better comparison. `geom_smooth` is a better method to display the data than `geom_line` because of our raw data. The plots are split with `facet_wrap` by their `Data_source`.

```{r 11-Plotting-generated-weather-with-ggplot-and-geom-smooth, fig.cap="Geom_smooth plot to compare observed and simulated temperature data.", message=FALSE}
ggplot(data=temperatures, aes(Date,Tmin)) +
  geom_smooth(aes(colour = factor(Year))) +
  facet_wrap(vars(Data_source)) +
  theme_bw(base_size = 20) +
  theme(legend.position = "none") +
  scale_x_date(date_labels = "%b")
```
\

## Calculate winter chill (in Chill Portions) for your synthetic weather, and illustrate your results as histograms and cumulative distributions.

We apply the `chilling` function to all of our `temperatures` that have the `Data_source` `"observed"` and `"simulated"`.

```{r 11 Chill calc and data prep, message=FALSE}
chill_observed<-chilling(
  stack_hourly_temps(
    temperatures[which(temperatures$Data_source=="observed"),],
    latitude = coords[2]),
  Start_JDay = 121, # Beginning May, start of "Winter" in southern hemisphere
  End_JDay = 243) # End of August, end of "Winter" in southern hemisphere

chill_simulated<-chilling(
  stack_hourly_temps(
    temperatures[which(temperatures$Data_source=="simulated"),],
    latitude = coords[2]),
  Start_JDay = 121,
  End_JDay = 243)

chill_comparison <- cbind(chill_observed ,Data_source="observed") 
chill_simulated <- cbind(chill_simulated ,Data_source="simulated")

chill_comparison <- rbind(chill_comparison, chill_simulated) #stacking simulated
#and observed

chill_comparison_full_seasons<-chill_comparison[
  which(chill_comparison$Perc_complete==100),] #only including complete years
```

For plotting the chill we use `geom_histogram`, the color is defined by the `Data_source`.
```{r chill-histo, fig.cap="A histogram that shows the chill portions for the observed and simulated data.", message=FALSE}
ggplot(chill_comparison_full_seasons, aes(x=Chill_portions)) + 
  geom_histogram(binwidth=1,aes(fill = factor(Data_source))) +
  theme_bw(base_size = 20) +
  labs(fill = "Data source") +
  xlab("Chill accumulation (Chill Portions)") +
  ylab("Frequency")
```

We can calculate and plot the Safe Winter Chill (SWC) as seen below. In 9 out of 10 years the amount of chilling can be expected to be above this value.
```{r 11-Cumulative-plotting-of-Chill-portions, fig.cap="The accumulated chill in Chill Portions. The red horizontal line marks the Safe Winter Chill."}
chill_simulations<-chill_comparison_full_seasons[
  which(chill_comparison_full_seasons$Data_source=="simulated"),]

ggplot(chill_simulations, aes(x=Chill_portions)) +
  stat_ecdf(geom = "step",lwd=1.5,col="blue") +
  ylab("Cumulative probability") +
  xlab("Chill accumulation (in Chill Portions)") +
  geom_hline(yintercept = 0.1, colour="red") +
  theme_bw(base_size = 20)
```
```{r 11 Calculating the SWC with quantile}
quantile(chill_simulations$Chill_portions, 0.1)
```
\

## Produce similar plots for the number of freezing hours (<0°C) in April (or October, if your site is in the Southern Hemisphere) for your location of interest.

First, I created a `freezing_hours` `step_model()`, that can be used as a model in the `tempResponse()` function. After that I applied it on the data in the same way as seen before, I also prepare the data in the same way and also filter all incomplete seasons.
```{r Creating freeze hours function and preparing the data}
freezing_hours_steps <- data.frame(lower=c(-1000,0), 
                                   upper=c(0,1000),
                                   weight=c(1,0))

freezing_hours <- function(x) step_model(x,freezing_hours_steps)

CapeTown_temperature <- cbind(weather_CapeTown[
  which(weather_CapeTown$Year %in% 1990:2005),])

observed_temperature <- stack_hourly_temps(CapeTown_temperature[,c("Year",
                                                                   "Month",
                                                                   "Day",
                                                                   "Tmin",
                                                                   "Tmax")],
                                           latitude = coords[2])$hourtemps

simulated_temperature <- stack_hourly_temps(generated_temp[,c("Year",
                                                              "Month",
                                                              "Day",
                                                              "Tmin",
                                                              "Tmax")], 
                                            latitude = coords[2])$hourtemps

freezing_observed <- tempResponse(make_JDay(observed_temperature),
                                  Start_JDay = 274,
                                  End_JDay = 304,
                                  models=list(Freezing_Hours=freezing_hours))

freezing_simulated <- tempResponse(make_JDay(simulated_temperature),
                                   Start_JDay = 274,
                                   End_JDay = 304,
                                   models=list(Freezing_Hours=freezing_hours))

freezing_comparison <- cbind(freezing_observed, Data_source="observed") 
freezing_simulated <- cbind(freezing_simulated, Data_source="simulated")


freezing_comparison <- rbind(freezing_comparison, freezing_simulated)

freezing_comparison_full_seasons<-freezing_comparison[
  which(freezing_comparison$Perc_complete==100),]
```

We can plot the data in the same way as seen before in Figure \@ref(fig:chill-histo). This time we can use the `freezing_comparison` dataset as input for the `ggplot` function.
```{r freeze-histo, fig.cap="A histogram that shows the freezing hours for the observed and simulated data."}
ggplot(freezing_comparison, aes(x=Freezing_Hours)) + 
  geom_histogram(binwidth=1,aes(fill = factor(Data_source))) +
  theme_bw(base_size = 20) +
  labs(fill = "Data source") +
  xlab("Freezing Hours") +
  ylab("Frequency")
```

My chosen location is near Cape Town, South Africa. As we can see in Figure \@ref(fig:freeze-histo), freezing hours are nearly non existent. Calculating something like a "Safe Freezing Hours" is not possible.
```{r 11-Cumulative-plotting-of-freezing-hours, fig.cap="The graph for accumulated freeze hours."}
freezing_simulations<-freezing_comparison_full_seasons[
  which(freezing_comparison_full_seasons$Data_source=="simulated"),]

ggplot(freezing_simulations, aes(x=Freezing_Hours)) +
  stat_ecdf(geom = "step",lwd=1.5,col="blue") +
  ylab("Cumulative probability") +
  xlab("Freeze hours accumulation") +
  theme_bw(base_size = 20)
```
