# Historic temperature scenarios

## For the location you chose for previous exercises, produce historic temperature scenarios representing several years of the historic record (your choice).
```{r 13-Loading-packages-and-data, echo=FALSE, message=FALSE}
require(chillR)
require(dplyr)
require(kableExtra)
coords <- c(18.8148, -33.9575)
weather_CapeTown <- read.csv("data/CapeTown_weather.csv")
```

After creating a baseline scenario for 2001, we can use the `temperature_scenario_from_records` function to create absolute scenarios for the years `c(1985, 1995, 2005, 2015)`. The absolute scenarios are adjusted to the baseline to be used in the `temperature_generation()` function.
```{r 13-Generating-4x100-years-of-weather-of-the-past, eval=FALSE}
require(chillR)
scenario_2001<-temperature_scenario_from_records(weather=weather_CapeTown,
                                                 year=2001)
scenario_2001

all_past_scenarios <- temperature_scenario_from_records(
  weather=weather_CapeTown,
  year=c(1985, 1995, 2005, 2015)) #creating absolute scenarios for those years

adjusted_scenarios <- temperature_scenario_baseline_adjustment(
  baseline=scenario_2001,
  temperature_scenario = all_past_scenarios) #transform to relative with bl 96

all_past_scenario_temps <- temperature_generation(
  weather=weather_CapeTown,
  years=c(1982,2020),
  sim_years=c(2001,2100),
  temperature_scenario = adjusted_scenarios)
```

```{r}
scenario_2001<-temperature_scenario_from_records(weather=weather_CapeTown,
                                                 year=2001)
scenario_2001

all_past_scenarios <- temperature_scenario_from_records(
  weather=weather_CapeTown,
  year=c(1985, 1995, 2005, 2015)) #creating absolute scenarios for those years

adjusted_scenarios <- temperature_scenario_baseline_adjustment(
  baseline=scenario_2001,
  temperature_scenario = all_past_scenarios) #transform to relative with bl 96
```


```{r, echo=FALSE}
t1 <- scenario_2001$`2001`$data
t2 <- all_past_scenarios$`1985`$data
t3 <- adjusted_scenarios$`1985`$data

t1 %>%
  kbl(escape = F, format = "html", caption = "Baseline 2001") %>%
  kable_paper("hover", full_width = FALSE, position = "float_left")
t2 %>%
  kbl(escape = F, format = "html", caption = "Absolute 1985") %>%
  kable_paper("hover", full_width = FALSE, position = "float_left")
t3 %>%
  kbl(escape = F, format = "html", caption = "Adjusted relative 1985") %>%
  kable_paper("hover", full_width = FALSE, position = "float_left")
```
<hr />

```{r 13-Saving-all- past-scenario-temperatures, echo=FALSE, eval=FALSE}
save_temperature_scenarios(all_past_scenario_temps,"data",
                                                   "past_scenario_temps")
```

```{r 13-Loading-all-past-scenarios, echo=FALSE}
all_past_scenario_temps <- load_temperature_scenarios("data",
                                                      "past_scenario_temps")
```

\

## Produce chill distributions for these scenarios and plot them.

```{r 13-Calculating chill-for-all-past-scenarios}
chill_hist_scenario_list <- tempResponse_daily_list(all_past_scenario_temps,
                                                  latitude=coords[2],
                                                  Start_JDay = 121,
                                                  End_JDay = 243)

scenarios<-names(chill_hist_scenario_list)[1:4]

all_scenarios<-chill_hist_scenario_list[[scenarios[1]]]
all_scenarios[,"scenario"] <- as.numeric(scenarios[1])


for (sc in scenarios[2:4])
 all_scenarios<-rbind(all_scenarios,
                      cbind(chill_hist_scenario_list[[sc]],scenario=as.numeric(sc)))

all_scenarios <- all_scenarios[which(all_scenarios$Perc_complete==100),]


actual_chill <- tempResponse_daily_list(weather_CapeTown,latitude=coords[2],
                                                  Start_JDay = 121,
                                                  End_JDay = 243)[[1]]
actual_chill <- actual_chill[which(actual_chill$Perc_complete==100),]
```


```{r 13-Plotting-the-past-chill-portions-in-a-violin-ggplot, fig.cap="Violin charts that display the chill accumulation for the simulated years of 1985,1995,2005,2015. The blue points represent the actual chill measured during this time"}
require(ggplot2)
ggplot(data=all_scenarios,aes(scenario,Chill_Portions,
                              fill=factor(scenario))) +
  geom_violin() +
  ylab("Chill accumulation (Chill Portions)") +
  xlab("Scenario year") +
  theme_bw(base_size=15) +
  ylim(c(0,90)) +
  geom_point(data=actual_chill,
             aes(End_year,Chill_Portions,fill="blue"),
             col="blue",show.legend = FALSE) +
  scale_fill_discrete(name="Scenario",
                      breaks = unique(all_scenarios$scenario)) 
```

