# Delineating temperature response phases with PLS regression

## Briefly explain why you shouldn’t take the results of a PLS regression analysis between temperature and phenology at face value. What do you need in addition in order to make sense of such outputs?

The phenology of trees is determined by more than just the factor of temperature. Nevertheless, we use temperature as the only input because it is easier to handle and sufficient for our purpose.
PLS is a method that is also used in machine learning. Without the necessary knowledge about the application behavior or without the right expectations, such a method can quickly lead to p-hacking or detect structures that are of no real importance.
We need clear expectations on the results, like in our case the detection of the chilling and forcing phase to use PLS in a meaningful way. Furthermore, detected structures should not be overestimated.


## Replicate the PLS analysis for the Roter Boskoop dataset that you used in a previous lesson.
```{r, warning=FALSE, message=FALSE}
library(chillR)
library(kableExtra)
Boskoop<-read_tab("data/Roter_Boskoop_bloom_1958_2019.csv")
Boskoop_first<-Boskoop[,1:2]
Boskoop_first[,"Year"]<-substr(Boskoop_first$First_bloom,1,4)
Boskoop_first[,"Month"]<-substr(Boskoop_first$First_bloom,5,6)
Boskoop_first[,"Day"]<-substr(Boskoop_first$First_bloom,7,8)
Boskoop_first<-make_JDay(Boskoop_first)
Boskoop_first<-Boskoop_first[,c("Pheno_year","JDay")]
colnames(Boskoop_first)<-c("Year","pheno")
```


```{r, warning=FALSE, message=FALSE}
kable(Boskoop_first[1:5,]) %>%
  kable_styling("striped", position = "left",font_size = 8)
```

```{r, warning=FALSE, message=FALSE}
KA_temps<-read_tab("data/TMaxTMin1958-2019_patched.csv")
KA_temps<-make_JDay(KA_temps)

kable(KA_temps[1:5,])  %>%
  kable_styling("striped", position = "left",font_size = 8)
```

```{r, warning=FALSE, message=FALSE}
PLS_results<-PLS_pheno(KA_temps,Boskoop_first)

kable(PLS_results$PLS_summary[1:5,])  %>%
  kable_styling("striped", position = "left",font_size = 8)
```


```{r, warning=FALSE, message=FALSE}
library(ggplot2)

PLS_gg<-PLS_results$PLS_summary
PLS_gg[,"Month"]<-trunc(PLS_gg$Date/100)
PLS_gg[,"Day"]<-PLS_gg$Date-PLS_gg$Month*100
PLS_gg[,"Date"]<-ISOdate(2002,PLS_gg$Month,PLS_gg$Day)
PLS_gg[which(PLS_gg$JDay<=0),"Date"]<-
  ISOdate(2001,
          PLS_gg$Month[which(PLS_gg$JDay<=0)],
          PLS_gg$Day[which(PLS_gg$JDay<=0)])
PLS_gg[,"VIP_importance"]<-PLS_gg$VIP>=0.8
PLS_gg[,"VIP_Coeff"]<-factor(sign(PLS_gg$Coef)*PLS_gg$VIP_importance)

VIP_plot<- ggplot(PLS_gg,aes(x=Date,y=VIP)) +
  geom_bar(stat='identity',aes(fill=VIP>0.8))
VIP_plot <- VIP_plot +
  scale_fill_manual(name="VIP", 
                    labels = c("<0.8", ">0.8"), 
                    values = c("FALSE"="grey", "TRUE"="blue")) +
  theme_bw(base_size=15) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank() )

```

```{r, warning=FALSE, message=FALSE}
coeff_plot<- ggplot(PLS_gg,aes(x=Date,y=Coef)) +
  geom_bar(stat='identity',aes(fill=VIP_Coeff)) +
  scale_fill_manual(name="Effect direction", 
                    labels = c("Advancing", "Unimportant","Delaying"), 
                    values = c("-1"="red", "0"="grey","1"="dark green")) +
  theme_bw(base_size=15) +
  ylab("PLS coefficient") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank() )

```

```{r, warning=FALSE, message=FALSE}
temp_plot<- ggplot(PLS_gg) +
  geom_ribbon(aes(x=Date,ymin=Tmean-Tstdev,ymax=Tmean+Tstdev),
              fill="grey") +
  geom_ribbon(aes(x=Date,ymin=Tmean-Tstdev*(VIP_Coeff==-1),
                  ymax=Tmean+Tstdev*(VIP_Coeff==-1)),
              fill="red") +
  geom_ribbon(aes(x=Date,ymin=Tmean-Tstdev*(VIP_Coeff==1),
                  ymax=Tmean+Tstdev*(VIP_Coeff==1)),
              fill="dark green") +
  geom_line(aes(x=Date,y=Tmean)) +
  theme_bw(base_size=15) +
  ylab(expression(paste(T[mean]," (°C)")))

```

```{r, warning=FALSE, message=FALSE}
library(patchwork)
library(plotly)



  plot<- (VIP_plot +
            coeff_plot +
            temp_plot +
            plot_layout(ncol=1,
                        guides = "collect")
          ) & theme(legend.position = "right",
                    legend.text = element_text(size=8),
                    legend.title = element_text(size=10),
                    axis.title.x=element_blank())

plot
```
\

## Write down your thoughts on why we’re not seeing the temperature response pattern we may have expected. What happened to the chill response?

In countries like Germany, the forcing period could be the stronger driver towards the flower development. It is also possible that PLS cannot detect the chill response, when there is no monotonic relationship between temperature and chill accumulation. A successful PLS regression depends on sufficient amounts of data. If the analysis is performed with not enough data, this can also lead to errors. However, this will probably not be the case in this example.

