# PLS regression with agroclimatic metrics

## Repeat the PLS_chill_force procedure for the ‘Roter Boskoop’ dataset. Include plots of daily chill and heat accumulation.
```{r}
library(chillR)
library(kableExtra)

temps<-read_tab("data/TMaxTMin1958-2019_patched.csv")
temps_hourly<-stack_hourly_temps(temps,latitude=50.6)

kable(head(temps))  %>%
  kable_styling("striped", position = "left",font_size = 8)

daychill<-daily_chill(hourtemps=temps_hourly,
            running_mean=1,
            models = list(Chilling_Hours = Chilling_Hours, Utah_Chill_Units = Utah_Model,
    Chill_Portions = Dynamic_Model, GDH = GDH))

Boskoop <- read_tab("data/Roter_Boskoop_bloom_1958_2019.csv")
Boskoop_first <- Boskoop[,1:2]
Boskoop_first[,"Year"] <- substr(Boskoop_first$First_bloom,1,4)
Boskoop_first[,"Month"] <- substr(Boskoop_first$First_bloom,5,6)
Boskoop_first[,"Day"] <- substr(Boskoop_first$First_bloom,7,8)
Boskoop_first <- make_JDay(Boskoop_first)
Boskoop_first <- Boskoop_first[,c("Pheno_year","JDay")]
colnames(Boskoop_first) <- c("Year","pheno")


plscf <- PLS_chill_force(daily_chill_obj=daychill,
                       bio_data_frame=Boskoop_first,
                       split_month=6,
                       chill_models = "Chill_Portions",
                       heat_models = "GDH",
                       runn_means = 11)
```


```{r}
kbl(head(plscf$Chill_Portions$GDH$PLS_summary), escape = F) %>%
  kable_paper("hover", full_width = F, position = "left")
```

```{r}
PLS_gg<-plscf$Chill_Portions$GDH$PLS_summary
PLS_gg[,"Month"] <- trunc(PLS_gg$Date/100)
PLS_gg[,"Day"] <- PLS_gg$Date-PLS_gg$Month*100
PLS_gg[,"Date"] <- ISOdate(2002,PLS_gg$Month,PLS_gg$Day)
PLS_gg[which(PLS_gg$JDay<=0),"Date"] <- 
  ISOdate(2001,
          PLS_gg$Month[which(PLS_gg$JDay<=0)],
          PLS_gg$Day[which(PLS_gg$JDay<=0)])
PLS_gg[,"VIP_importance"] <- PLS_gg$VIP>=0.8
PLS_gg[,"VIP_Coeff"] <- factor(sign(PLS_gg$Coef)*PLS_gg$VIP_importance)

chill_start_JDay<--48 #!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
chill_end_JDay<-72
heat_start_JDay<--8
heat_end_JDay<-110

chill_start_date<-ISOdate(2001,12,31)+chill_start_JDay*24*3600
chill_end_date<-ISOdate(2001,12,31)+chill_end_JDay*24*3600
heat_start_date<-ISOdate(2001,12,31)+heat_start_JDay*24*3600
heat_end_date<-ISOdate(2001,12,31)+heat_end_JDay*24*3600
```

```{r}
library(ggplot2)

temp_plot<- ggplot(PLS_gg,x=Date) +
  annotate("rect",
           xmin = chill_start_date,
           xmax = chill_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "blue") +
  annotate("rect",
           xmin = heat_start_date,
           xmax = heat_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "red") +
  annotate("rect",
           xmin = ISOdate(2001,12,31) +
             min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           xmax = ISOdate(2001,12,31) +
             max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "black") +
  geom_vline(xintercept = ISOdate(2001,12,31) +
               median(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             linetype = "dashed") +
  geom_ribbon(aes(x=Date,
                  ymin=MetricMean - MetricStdev ,
                  ymax=MetricMean + MetricStdev ),
              fill="grey") +
  geom_ribbon(aes(x=Date,
                  ymin=MetricMean - MetricStdev * (VIP_Coeff==-1),
                  ymax=MetricMean + MetricStdev * (VIP_Coeff==-1)),
              fill="red") +
  geom_ribbon(aes(x=Date,
                  ymin=MetricMean - MetricStdev * (VIP_Coeff==1),
                  ymax=MetricMean + MetricStdev * (VIP_Coeff==1)),
              fill="dark green") +
  geom_line(aes(x=Date,y=MetricMean ))
```

```{r}
temp_plot<- temp_plot +
  facet_wrap(vars(Type), scales = "free_y",
             strip.position="left",
             labeller = labeller(Type = as_labeller(
               c(Chill="Chill (CP)",Heat="Heat (GDH)")))) +
  ggtitle("Daily chill and heat accumulation rates") +
  theme_bw(base_size=15) + 
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y = element_text(size =12),
        plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()
        )

temp_plot
```

```{r}
VIP_plot<- ggplot(PLS_gg,aes(x=Date,y=VIP)) +
  annotate("rect",
           xmin = chill_start_date,
           xmax = chill_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "blue") +
  annotate("rect",
           xmin = heat_start_date,
           xmax = heat_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "red") +
  annotate("rect",
           xmin = ISOdate(2001,12,31) +
             min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           xmax = ISOdate(2001,12,31) +
             max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "black") +
  geom_vline(xintercept = ISOdate(2001,12,31) +
               median(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             linetype = "dashed") +
  geom_bar(stat='identity',aes(fill=VIP>0.8))
```

```{r}
VIP_plot <- VIP_plot + facet_wrap(vars(Type), scales="free",
             strip.position="left",
             labeller = labeller(Type = as_labeller(
               c(Chill="VIP for chill",Heat="VIP for heat")))) +
  scale_y_continuous(
    limits=c(0,max(plscf$Chill_Portions$GDH$PLS_summary$VIP))) +
  ggtitle("Variable Importance in the Projection (VIP) scores") +
  theme_bw(base_size=15) + 
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y = element_text(size =12),
        plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()
        )
```

```{r}
VIP_plot <- VIP_plot +
  scale_fill_manual(name="VIP", 
                    labels = c("<0.8", ">0.8"), 
                    values = c("FALSE"="grey", "TRUE"="blue")) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

VIP_plot
```

```{r}
coeff_plot<- ggplot(PLS_gg,aes(x=Date,y=Coef)) +
  annotate("rect",
           xmin = chill_start_date,
           xmax = chill_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "blue") +
  annotate("rect",
           xmin = heat_start_date,
           xmax = heat_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "red") +
  annotate("rect",
           xmin = ISOdate(2001,12,31) +
             min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           xmax = ISOdate(2001,12,31) +
             max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "black") +
  geom_vline(xintercept = ISOdate(2001,12,31) +
               median(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             linetype = "dashed") +
  geom_bar(stat='identity',aes(fill=VIP_Coeff))
```

```{r}
coeff_plot <- coeff_plot + facet_wrap(vars(Type), scales="free",
             strip.position="left",
             labeller = labeller(
               Type = as_labeller(
                 c(Chill="MC for chill",Heat="MC for heat")))) +
  scale_y_continuous(
    limits=c(min(plscf$Chill_Portions$GDH$PLS_summary$Coef),
             max(plscf$Chill_Portions$GDH$PLS_summary$Coef))) +
  ggtitle("Model coefficients (MC)") +
  theme_bw(base_size=15) + 
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y = element_text(size =12),
        plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()
        )
```

```{r, message=FALSE, warning=FALSE}
coeff_plot <- coeff_plot + facet_wrap(vars(Type), scales="free",
             strip.position="left",
             labeller = labeller(
               Type = as_labeller(
                 c(Chill="MC for chill",Heat="MC for heat")))) +
  scale_y_continuous(
    limits=c(min(plscf$Chill_Portions$GDH$PLS_summary$Coef),
             max(plscf$Chill_Portions$GDH$PLS_summary$Coef))) +
  ggtitle("Model coefficients (MC)") +
  theme_bw(base_size=15) + 
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y = element_text(size =12),
        plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()
        )
```

```{r, message=FALSE, warning=FALSE}
coeff_plot <- coeff_plot +  scale_fill_manual(name="Effect direction", 
                    labels = c("Advancing", "Unimportant","Delaying"), 
                    values = c("-1"="red", "0"="grey","1"="dark green")) +
  ylab("PLS coefficient") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

coeff_plot
```

```{r, message=FALSE, warning=FALSE}
library(patchwork)

plot<- (VIP_plot +
          coeff_plot +
          temp_plot +
          plot_layout(ncol=1,
            guides = "collect")
        ) & theme(legend.position = "right",
                  legend.text = element_text(size=8),
                  legend.title = element_text(size=10),
                  axis.title.x=element_blank())

plot
```

```{r, message=FALSE, warning=FALSE}
plot_PLS_chill_force<-function(plscf,
                               chill_metric="Chill_Portions",
                               heat_metric="GDH",
                               chill_label="CP",
                               heat_label="GDH",
                               chill_phase=c(-48,62),
                               heat_phase=c(-5,105.5),
                               cultivar_name=" ")
{
  PLS_gg<-plscf[[chill_metric]][[heat_metric]]$PLS_summary
  PLS_gg[,"Month"]<-trunc(PLS_gg$Date/100)
  PLS_gg[,"Day"]<-PLS_gg$Date-PLS_gg$Month*100
  PLS_gg[,"Date"]<-ISOdate(2002,PLS_gg$Month,PLS_gg$Day)
  PLS_gg[which(PLS_gg$JDay<=0),"Date"]<-ISOdate(2001,PLS_gg$Month[which(PLS_gg$JDay<=0)],PLS_gg$Day[which(PLS_gg$JDay<=0)])
  PLS_gg[,"VIP_importance"]<-PLS_gg$VIP>=0.8
  PLS_gg[,"VIP_Coeff"]<-factor(sign(PLS_gg$Coef)*PLS_gg$VIP_importance)
  
  chill_start_date<-ISOdate(2001,12,31)+chill_phase[1]*24*3600
  chill_end_date<-ISOdate(2001,12,31)+chill_phase[2]*24*3600
  heat_start_date<-ISOdate(2001,12,31)+heat_phase[1]*24*3600
  heat_end_date<-ISOdate(2001,12,31)+heat_phase[2]*24*3600

  
  temp_plot<- ggplot(PLS_gg) +
    annotate("rect",
             xmin = chill_start_date,
             xmax = chill_end_date,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "blue") +
    annotate("rect",
             xmin = heat_start_date,
             xmax = heat_end_date,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "red") +
    annotate("rect",
             xmin = ISOdate(2001,12,31) + min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             xmax = ISOdate(2001,12,31) + max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "black") +
    geom_vline(xintercept = ISOdate(2001,12,31) + median(plscf$pheno$pheno,na.rm=TRUE)*24*3600, linetype = "dashed") +
    geom_ribbon(aes(x=Date,
                    ymin=MetricMean - MetricStdev ,
                    ymax=MetricMean + MetricStdev ),
                fill="grey") +
    geom_ribbon(aes(x=Date,
                    ymin=MetricMean - MetricStdev * (VIP_Coeff==-1),
                    ymax=MetricMean + MetricStdev * (VIP_Coeff==-1)),
                fill="red") +
    geom_ribbon(aes(x=Date,
                    ymin=MetricMean - MetricStdev * (VIP_Coeff==1),
                    ymax=MetricMean + MetricStdev * (VIP_Coeff==1)),
                fill="dark green") +
    geom_line(aes(x=Date,y=MetricMean )) +
    facet_wrap(vars(Type), scales = "free_y",
               strip.position="left",
               labeller = labeller(Type = as_labeller(c(Chill=paste0("Chill (",chill_label,")"),Heat=paste0("Heat (",heat_label,")")) )) ) +
    ggtitle("Daily chill and heat accumulation rates") +
    theme_bw(base_size=15) + 
    theme(strip.background = element_blank(),
          strip.placement = "outside",
          strip.text.y = element_text(size =12),
          plot.title = element_text(hjust = 0.5),
          axis.title.y=element_blank()
          )
  
  VIP_plot<- ggplot(PLS_gg,aes(x=Date,y=VIP)) +
    annotate("rect",
             xmin = chill_start_date,
             xmax = chill_end_date,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "blue") +
    annotate("rect",
             xmin = heat_start_date,
             xmax = heat_end_date,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "red") +
    annotate("rect",
             xmin = ISOdate(2001,12,31) + min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             xmax = ISOdate(2001,12,31) + max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "black") +
    geom_vline(xintercept = ISOdate(2001,12,31) + median(plscf$pheno$pheno,na.rm=TRUE)*24*3600, linetype = "dashed") +
    geom_bar(stat='identity',aes(fill=VIP>0.8)) +
    facet_wrap(vars(Type), scales="free",
               strip.position="left",
               labeller = labeller(Type = as_labeller(c(Chill="VIP for chill",Heat="VIP for heat") )) ) +
    scale_y_continuous(limits=c(0,max(plscf[[chill_metric]][[heat_metric]]$PLS_summary$VIP))) +
    ggtitle("Variable Importance in the Projection (VIP) scores") +
    theme_bw(base_size=15) + 
    theme(strip.background = element_blank(),
          strip.placement = "outside",
          strip.text.y = element_text(size =12),
          plot.title = element_text(hjust = 0.5),
          axis.title.y=element_blank()
          ) +
    scale_fill_manual(name="VIP", 
                      labels = c("<0.8", ">0.8"), 
                      values = c("FALSE"="grey", "TRUE"="blue")) +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank())
  
  coeff_plot<- ggplot(PLS_gg,aes(x=Date,y=Coef)) +
    annotate("rect",
             xmin = chill_start_date,
             xmax = chill_end_date,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "blue") +
    annotate("rect",
             xmin = heat_start_date,
             xmax = heat_end_date,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "red") +
    annotate("rect",
             xmin = ISOdate(2001,12,31) + min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             xmax = ISOdate(2001,12,31) + max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             ymin = -Inf,
             ymax = Inf,
             alpha = .1,fill = "black") +
    geom_vline(xintercept = ISOdate(2001,12,31) + median(plscf$pheno$pheno,na.rm=TRUE)*24*3600, linetype = "dashed") +
    geom_bar(stat='identity',aes(fill=VIP_Coeff)) +
    facet_wrap(vars(Type), scales="free",
               strip.position="left",
               labeller = labeller(Type = as_labeller(c(Chill="MC for chill",Heat="MC for heat") )) ) +
    scale_y_continuous(limits=c(min(plscf[[chill_metric]][[heat_metric]]$PLS_summary$Coef),
                                max(plscf[[chill_metric]][[heat_metric]]$PLS_summary$Coef))) +
    ggtitle("Model coefficients (MC)") +
    theme_bw(base_size=15) + 
    theme(strip.background = element_blank(),
          strip.placement = "outside",
          strip.text.y = element_text(size =12),
          plot.title = element_text(hjust = 0.5),
          axis.title.y=element_blank()
          ) +
    scale_fill_manual(name="Effect direction", 
                      labels = c("Advancing", "Unimportant","Delaying"), 
                      values = c("-1"="red", "0"="grey","1"="dark green")) +
    ylab("PLS coefficient") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank())
  
  library(patchwork)
  
  plot<- (VIP_plot +
            coeff_plot +
            temp_plot +
            plot_layout(ncol=1,
                        guides = "collect")) +
          theme(legend.position = "right",
                    legend.text = element_text(size=8),
                    legend.title = element_text(size=10),
                    axis.title.x=element_blank()) + 
          plot_annotation(title = cultivar_name)

plot

}
```
\

## Run PLS_chill_force analyses for all three major chill models. Delineate your best estimates of chilling and forcing phases for all of them.

```{r, message=FALSE, warning=FALSE}
daychill<-daily_chill(hourtemps=temps_hourly,
            running_mean=9,
            models = list(Chilling_Hours = Chilling_Hours, Utah_Chill_Units = Utah_Model,
    Chill_Portions = Dynamic_Model, GDH = GDH)
    )

plscf<-PLS_chill_force(daily_chill_obj=daychill,
                       bio_data_frame=Boskoop_first,
                       split_month=6,
                       chill_models = c("Chilling_Hours", "Utah_Chill_Units", "Chill_Portions"),
                       heat_models = c("GDH"))
```

We can delineate the estimates of the chilling and forcing phases by having a direct look on the output of the PLS analysis another possibility is to estimate the phases with the help of the plotted version. `options(max.print=636)` allows to print the whole dataset, which had in this case 636 rows. The Coef and the VIP column contain the data that also the function uses for the plotting later. I prefer to use the plotted version of the data to delineate the phases, because this gives me a better overview over the longer development.

```{r, eval=FALSE}
options(max.print=636)
plscf
```
My estimated delineations are included in the table below. I wasnt able to delineate the chilling phase for the Chilling Hours model, because no model coefficients seemed to have an advancing effect. For both other models I was able to delineate the chilling period. For the GDH model, I set the delineation estimates once and did not change them, even though I observed slight changes in the advancing effects between the different chill models.

```{r, echo=FALSE}
name <- c("Chilling Hours", "Utah Model", "Dynamic Model")
start_c <- c(0,-13,24)
end_c <- c(0,65,62)
start_h <- c(8,8,8)
end_h <- c(106,106,106)
delineations <- data.frame(name, start_c, end_c, start_h, end_h, row.names = )

kbl(delineations, col.names = NULL) %>%
  kable_paper("hover", full_width = F, position = "left") %>%
  add_header_above(c(" ", "Start", "End", "Start", "End")) %>%
  add_header_above(c(" ", "Chill" = 2, "Heat" = 2))
```



\

## Plot results for all three analyses, including shaded plot areas for the chilling and forcing periods you estimated.

Below I plotted the estimated delineations together with the previously created function. For the Chilling Hours plot (as already mentioned above) no chilling period can be determined
```{r}
plot_PLS_chill_force(plscf,
                     chill_metric="Chilling_Hours",
                     heat_metric="GDH",
                     chill_label="CH",
                     heat_label="GDH",
                     chill_phase=c(0,0),
                     heat_phase=c(8,106))
```

```{r}
plot_PLS_chill_force(plscf,
                     chill_metric="Utah_Chill_Units",
                     heat_metric="GDH",
                     chill_label="CU",
                     heat_label="GDH",
                     chill_phase=c(-13,65),
                     heat_phase=c(8,108))
```

```{r}
plot_PLS_chill_force(plscf,
                     chill_metric="Chill_Portions",
                     heat_metric="GDH",
                     chill_label="CP",
                     heat_label="GDH",
                     chill_phase=c(-24,62),
                     heat_phase=c(8,108))
```





