---
bibliography: libs/references.bib
link-citations: TRUE
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

# Chapter 20 PLS regression with agroclimatic metrics
\
**1. Repeat the PLS_chill_force procedure for the ‘Roter Boskoop’ dataset. Include plots of daily chill and heat accumulation.**
```{r}
library(chillR)
library(kableExtra)

temps<-read_tab("data/TMaxTMin1958-2019_patched.csv")
temps_hourly<-stack_hourly_temps(temps,latitude=50.6)

kable(head(temps))  %>%
  kable_styling("striped", position = "left",font_size = 8)

daychill<-daily_chill(hourtemps=temps_hourly,
            running_mean=1,
            models = list(Chilling_Hours = Chilling_Hours, Utah_Chill_Units = Utah_Model,
    Chill_Portions = Dynamic_Model, GDH = GDH))

Boskoop<-read_tab("data/Roter_Boskoop_bloom_1958_2019.csv")
Boskoop_first<-Boskoop[,1:2]
Boskoop_first[,"Year"]<-substr(Boskoop_first$First_bloom,1,4)
Boskoop_first[,"Month"]<-substr(Boskoop_first$First_bloom,5,6)
Boskoop_first[,"Day"]<-substr(Boskoop_first$First_bloom,7,8)
Boskoop_first<-make_JDay(Boskoop_first)
Boskoop_first<-Boskoop_first[,c("Pheno_year","JDay")]
colnames(Boskoop_first)<-c("Year","pheno")


plscf<-PLS_chill_force(daily_chill_obj=daychill,
                       bio_data_frame=Boskoop_first,
                       split_month=6,
                       chill_models = "Chill_Portions",
                       heat_models = "GDH",
                       runn_means = 11)

kable(head(plscf$Chill_Portions$GDH$PLS_summary))  %>%
  kable_styling("striped", position = "left",font_size = 10)
```

```{r}
PLS_gg<-plscf$Chill_Portions$GDH$PLS_summary
PLS_gg[,"Month"]<-trunc(PLS_gg$Date/100)
PLS_gg[,"Day"]<-PLS_gg$Date-PLS_gg$Month*100
PLS_gg[,"Date"]<-ISOdate(2002,PLS_gg$Month,PLS_gg$Day)
PLS_gg[which(PLS_gg$JDay<=0),"Date"]<-
  ISOdate(2001,
          PLS_gg$Month[which(PLS_gg$JDay<=0)],
          PLS_gg$Day[which(PLS_gg$JDay<=0)])
PLS_gg[,"VIP_importance"]<-PLS_gg$VIP>=0.8
PLS_gg[,"VIP_Coeff"]<-factor(sign(PLS_gg$Coef)*PLS_gg$VIP_importance)

chill_start_JDay<--48 #!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
chill_end_JDay<-72
heat_start_JDay<--8
heat_end_JDay<-110

chill_start_date<-ISOdate(2001,12,31)+chill_start_JDay*24*3600
chill_end_date<-ISOdate(2001,12,31)+chill_end_JDay*24*3600
heat_start_date<-ISOdate(2001,12,31)+heat_start_JDay*24*3600
heat_end_date<-ISOdate(2001,12,31)+heat_end_JDay*24*3600
```

```{r}
library(ggplot2)

temp_plot<- ggplot(PLS_gg,x=Date) +
  annotate("rect",
           xmin = chill_start_date,
           xmax = chill_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "blue") +
  annotate("rect",
           xmin = heat_start_date,
           xmax = heat_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "red") +
  annotate("rect",
           xmin = ISOdate(2001,12,31) +
             min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           xmax = ISOdate(2001,12,31) +
             max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "black") +
  geom_vline(xintercept = ISOdate(2001,12,31) +
               median(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             linetype = "dashed") +
  geom_ribbon(aes(x=Date,
                  ymin=MetricMean - MetricStdev ,
                  ymax=MetricMean + MetricStdev ),
              fill="grey") +
  geom_ribbon(aes(x=Date,
                  ymin=MetricMean - MetricStdev * (VIP_Coeff==-1),
                  ymax=MetricMean + MetricStdev * (VIP_Coeff==-1)),
              fill="red") +
  geom_ribbon(aes(x=Date,
                  ymin=MetricMean - MetricStdev * (VIP_Coeff==1),
                  ymax=MetricMean + MetricStdev * (VIP_Coeff==1)),
              fill="dark green") +
  geom_line(aes(x=Date,y=MetricMean ))

temp_plot
```

```{r}
temp_plot<- temp_plot +
  facet_wrap(vars(Type), scales = "free_y",
             strip.position="left",
             labeller = labeller(Type = as_labeller(
               c(Chill="Chill (CP)",Heat="Heat (GDH)")))) +
  ggtitle("Daily chill and heat accumulation rates") +
  theme_bw(base_size=15) + 
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y = element_text(size =12),
        plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()
        )

temp_plot
```

```{r}
VIP_plot<- ggplot(PLS_gg,aes(x=Date,y=VIP)) +
  annotate("rect",
           xmin = chill_start_date,
           xmax = chill_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "blue") +
  annotate("rect",
           xmin = heat_start_date,
           xmax = heat_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "red") +
  annotate("rect",
           xmin = ISOdate(2001,12,31) +
             min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           xmax = ISOdate(2001,12,31) +
             max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "black") +
  geom_vline(xintercept = ISOdate(2001,12,31) +
               median(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             linetype = "dashed") +
  geom_bar(stat='identity',aes(fill=VIP>0.8))

VIP_plot
```

```{r}
VIP_plot <- VIP_plot + facet_wrap(vars(Type), scales="free",
             strip.position="left",
             labeller = labeller(Type = as_labeller(
               c(Chill="VIP for chill",Heat="VIP for heat")))) +
  scale_y_continuous(
    limits=c(0,max(plscf$Chill_Portions$GDH$PLS_summary$VIP))) +
  ggtitle("Variable Importance in the Projection (VIP) scores") +
  theme_bw(base_size=15) + 
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y = element_text(size =12),
        plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()
        )
  
VIP_plot
```

```{r}
VIP_plot <- VIP_plot +
  scale_fill_manual(name="VIP", 
                    labels = c("<0.8", ">0.8"), 
                    values = c("FALSE"="grey", "TRUE"="blue")) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

VIP_plot
```

```{r}
coeff_plot<- ggplot(PLS_gg,aes(x=Date,y=Coef)) +
  annotate("rect",
           xmin = chill_start_date,
           xmax = chill_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "blue") +
  annotate("rect",
           xmin = heat_start_date,
           xmax = heat_end_date,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "red") +
  annotate("rect",
           xmin = ISOdate(2001,12,31) +
             min(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           xmax = ISOdate(2001,12,31) +
             max(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
           ymin = -Inf,
           ymax = Inf,
           alpha = .1,fill = "black") +
  geom_vline(xintercept = ISOdate(2001,12,31) +
               median(plscf$pheno$pheno,na.rm=TRUE)*24*3600,
             linetype = "dashed") +
  geom_bar(stat='identity',aes(fill=VIP_Coeff))

coeff_plot
```

```{r}
coeff_plot <- coeff_plot + facet_wrap(vars(Type), scales="free",
             strip.position="left",
             labeller = labeller(
               Type = as_labeller(
                 c(Chill="MC for chill",Heat="MC for heat")))) +
  scale_y_continuous(
    limits=c(min(plscf$Chill_Portions$GDH$PLS_summary$Coef),
             max(plscf$Chill_Portions$GDH$PLS_summary$Coef))) +
  ggtitle("Model coefficients (MC)") +
  theme_bw(base_size=15) + 
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y = element_text(size =12),
        plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()
        )
  
coeff_plot 
```

```{r}
coeff_plot <- coeff_plot + facet_wrap(vars(Type), scales="free",
             strip.position="left",
             labeller = labeller(
               Type = as_labeller(
                 c(Chill="MC for chill",Heat="MC for heat")))) +
  scale_y_continuous(
    limits=c(min(plscf$Chill_Portions$GDH$PLS_summary$Coef),
             max(plscf$Chill_Portions$GDH$PLS_summary$Coef))) +
  ggtitle("Model coefficients (MC)") +
  theme_bw(base_size=15) + 
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y = element_text(size =12),
        plot.title = element_text(hjust = 0.5),
        axis.title.y=element_blank()
        )
  
coeff_plot 
```

```{r}
coeff_plot <- coeff_plot +  scale_fill_manual(name="Effect direction", 
                    labels = c("Advancing", "Unimportant","Delaying"), 
                    values = c("-1"="red", "0"="grey","1"="dark green")) +
  ylab("PLS coefficient") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())

coeff_plot
```

```{r}
library(patchwork)

plot<- (VIP_plot +
          coeff_plot +
          temp_plot +
          plot_layout(ncol=1,
            guides = "collect")
        ) & theme(legend.position = "right",
                  legend.text = element_text(size=8),
                  legend.title = element_text(size=10),
                  axis.title.x=element_blank())

plot
```

\
**2. Run PLS_chill_force analyses for all three major chill models. Delineate your best estimates of chilling and forcing phases for all of them.**

\
**3. Plot results for all three analyses, including shaded plot areas for the chilling and forcing periods you estimated.**

